### 2020-10-14 papers/sugar-synth/main.tex

\documentclass{llncs}
%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
% \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

% \documentclass[conference,10pt]{IEEEtran}

% \bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
% \citestyle{acmauthoryear}   %% For author/year citations

\input{macro.tex}

\title{Automated inference of production rules for glycans}
\author{Ashutosh Gupta and Mukund Thattai and Ansuman Biswas}
%\institute{IITB and NCBS}

\date{May 2019}

\begin{document}

\maketitle

\begin{abstract}
  Glycans are tree-like polymers made up of sugar monomer building blocks. They are found on the surface of all living cells, and distinct glycan trees act as identity markers for distinct cell types. Proteins called GTase enzymes assemble glycans via the successive addition of monomer building blocks. The rules by which these enzymes operate are not fully understood. In this paper, we present a novel SMT-solver-based iterative method that infers the assembly process of the glycans by analyzing the set of glycans found in a cell. We have build a tool based on this method and have used it to infer rules based on published glycan data sets.
\end{abstract}

\section{Introduction}
\label{sec:intro}
\input{intro}

\section{Production of Glycans}
\label{sec:bio}
\input{bio.tex}

% \clearpage

\section{Motivating example}
\label{sec:motivation}
\input{motivation.tex}

\section{Modelling of the synthesis problem}
\label{sec:model}
\input{model.tex}

% \clearpage

\section{Method for the synthesis problem}
\label{sec:algo}
\input{algo.tex}

\section{Variations of the synthesis problem}
\label{sec:variations}
\input{variations.tex}

\section{Experiments}
\label{sec:experiments}
\input{experiments.tex}

\section{Related work}
\label{sec:related}
\input{related.tex}

\section{Conclusion}
% ##limitations in the model because of limited information,
% a)the boundary between two compartments is fuzzy, 
% b) the growths of two branches are correlated/ not correlated..and some more points like these, for which wealth of data is not available, hence the model was unable to incorporate the specifics
\label{sec:conclusion}
\input{conclusion.tex}

\bibliographystyle{unsrt}
\bibliography{biblio}

\appendix

\section{Implementation optimizations}

% \subsection{Optimizations}
We have incorporated several optimizations to gain efficiency or reduce
the number of iterations in the method. Let us discuss some of the remarkable
optimizations.

\paragraph{Ordered templates}
Our method constructs a list of templates.
Since the templates are symmetric, there is a potential of unnecessary
iterations.
Let us suppose the method rejects a set of learned rules.
In the subsequent queries include the constraints that reject the set of rules,
the solvers may choose the same set of rules again by permuting the mapping
from the templates to the learned rule.
To avoid these iterations, we define an arbitrary total order over the rules.
We add ordering constraints stating that any assignments to the list of templates should produce
ordered rules with respect to the total order.

\paragraph{Quantifier instantiations}
The constraints for rejecting the counterexample molecule is universally quantified at line~\ref{line:negCons} in algorithm~\ref{alg:sugar-synth}.
For any assignment, the solver needs to try sufficiently many instantiations of the quantifiers
to ensure that the query is not unsatisfiable.
The instantiations sufficiently slow down the solving process.
We assist the solver by also adding an instantiation of the quantifiers that is equal
to the values of the corresponding variables in the assignment $a$ at line~\ref{line:negModel}.
We observe that for some inputs only adding the instantiations and not the universally quantified
formula is more effective.

\paragraph{Constraints for counterexample molecules}
In the presentation of \textsc{SugarSynth}, we construct constraints of counterexample molecules
at line~\ref{line:consNewR} in each iteration.
The repeated work is impractical because the formula management system of Z3 will be overwhelmed by
the construction of many terms repeatedly. In our implementation, we construct the constraints
once outside the while loop.
We pass the templates as the second parameter instead of concrete rules to~\textsc{EncodeProduce}.
Later at the solving time in line~\ref{line:negModel}, constraints are added to assign values
of the template variables that were returned by the solver at line~\ref{line:posModel}.
Due to the assignments, the templates become concrete rules in the context of solving.

\section{Features to support in future}

\subsection{Full organism data vs single cell data}

The experiments that observe the set of glycan molecules are of two kinds. In one kind, we isolate a single cell type of an organism and identify all the glycans present in the cells. In our presentation, we have assumed this kind of source of data. However, the experiments are difficult to conduct.
In another and more convenient way, we smash the whole organism and identify all the glycans present in all the cells.
So the information that which glycans are coming from which
cells is unknown.
In this situation, the synthesis has another task to map the molecules
in $\mu$ to several cell types.
A glycan may be present in multiple cell types.

Our method is easily adapted to consider this kind of data.
We search for a small covering set of subsets of
$\powerset{\mu}$ such that each set in the cover allows a small
set of production rules.
The cover sets indicate the different types of cells in terms of
the presence of glycans.
This variation makes the problem particularly hard.
There can be a potentially large number of covering subsets.
So far, we have not encountered a large enough data set such that
we apply the variation effectively.

\subsection{Incomplete data}

The experiments are imperfect. They may not detect all possible glycans
in a cell.
We may need to leave the possibility of allowing a few more molecules to be
producible beyond $\mu$.
We replace the no extra molecule constraint by a constraint that allow molecules
that are `similar' to molecules in $\mu$.
However, we could not imagine any measure of similarity that naturally stems from
biological intuition.
For now, our tool supports the count of the number of monomer differences
as a measure of similarity.

\section{Extended related work}

% Success of formal methods
Formal methods have successful in applications to a vast range of
problems from the analysis of systems.
%
In particular,
the verification of hardware designs
using SAT solver based model-checking has been applied
to industrial-scale problems~\cite{biere1999symbolic2}.
%
The verification of software systems is a much harder problem, and
many methods like bounded model-checking~\cite{biere2003bounded},
abstract interpretation~\cite{lattice77}, and
counterexample guided abstraction refinement (CEGAR)~\cite{ClarkeCEGAR} have
been successful for the problem.
%

Synthesis of programs that implements a given behavior
has been a focus of research for a while~\cite{PnueliSynthesis}.
%
In recent years, the synthesis of programs using SAT/SMT solvers
has gained momentum.
%
The approach encodes the search of a program that exhibits a certain
behavior into a satisfiability problem.
%
The solvers attempt to find a solution of the satisfiability problem.
%
The solution is the synthesized program.
%
In~\cite{SrivastavaSynthesis,Solar-Lezama2005},
the method has been successfully applied
to find programs that satisfy quantified specifications like sorting
programs, and have missing ``holes'' and need proper implementations
for them.
%
In~\cite{exampleSynth},
a set of pairs of input and output examples is the specification of a program,
and the synthesis method searches for programs in a space defined by a
template~\cite{sygus}.

% Boolean networks

% Program synthesis
Since the formal methods have been exhibiting effectiveness in the vast
range of problems, the
systems biology community has been applying the methods for various
biological problems~\cite{fisher2007executable}.
The formal methods approach is distinctly different from the statistical
approach traditionally used in biology.
In formal methods, we aim to match precisely the expected behavior rather
than learn approximate artifacts from biological data.
The use of formal methods belongs to two broad categories:
the analysis of biological models and the synthesis of the models.

The key focus has been to model gene regulatory networks (GRNs), since
they are the core of the central dogma of biology.
Boolean networks are often used to model
GRNs to find stable points or attractors, i.e., nodes in the networks
where a system eventually reaches no matter where it starts~\cite{wang2012BooleanOverview}.
% In~\cite{boolean-networks-stabilty}, the networks are used to evaluate
% stability of GRNs under mutations.
The Boolean networks are not often sufficient to adequately model the behavior
of GRNs.
The continuous-time Markov chains (CTMCs) are used to bring in the aspects of
timing and probabilistic constraints to GRNs.
The transient behaviors of GRNs are also crucial for various aspects
of biology.
% For example, the decision making in some cells is based on
% response timing to stimulus~\cite{response-timing-paper}.
In~\cite{delayedCTMC}, a methods based on
model-checking is used to estimate the time evolution of the probability
distributions over states of GRNs.

There are many biological processes where we can observe the system behavior,
but the exact mechanisms of the processes are not known.
%
Recently, the methods for formal synthesis are finding their applications in biology
\cite{dunn2014defining,xuPluripotency,booleanModelKarp13,paoletti2014analyzing}.
%
In~\cite{koksal2013synthesis},
GRNs with unknown interactions are modeled using
a automata-like model with missing components
and the specification is the variations of the behavior of the system under mutations.
Their approach uses a counterexample-guided inductive synthesis (CEGIS) based algorithm~\cite{cegis}.
CEGIS is a framework for synthesis. It first finds a program that may satisfy `samples' from
the specification. If the synthesized program satisfies the specifications, the CEGIS terminates.
Otherwise, the method learns a new sample where the program violates the specifications.
It adds the sample to the set and goes to the next iteration.
Typically, the constraint solvers find the programs during intermediate steps of CEGIS by
solving the generated queries.
%
Our method follows a similar pattern, where
a set of sampled observations is a specification, i.e., output molecules,
and we need to find the governing programs, i.e., production rules.
%
In a subsequent work~\cite{fisher2015synthesising}, a Boolean network model is used,
the functions attached to the nodes are considered unknown, and
a different kind of data sets provides the behavior specification.
In this approach, their earlier method is adopted to work in the new situation.
%

As far as we know, there has been no similar computation based analysis of glycan production rules.
%
However, there has been a theoretical analysis of the properties of the production
rules in~\cite{Jaiman2018}. The work identifies the conditions for the production
of a finite set of molecules and the cases of unique production methods.
%
In our work, we are taking the computational approach of synthesis from formal methods
instead of looking for the theoretical conditions.

% 
% To resolve this paradox we borrow from the field of
% al-gorithmic self-assembly, which explores how small buildingblocks
% with stochastic local interactions assemble into various global
% configurations [23]–[25].  A central inverse problemin self-assembly
% is to design building blocks that assembleinto a target shape.  This
% process has been studied using thetheoretical framework of Wang tiles:
% square building blockswith colored sides that stick to one another
% along sides withmatching colors [26].  For efficient assembly of Wang
% tiles,their coloring must be cleverly chosen so the target shape isthe
% unique terminal output of the stochastic assembly process[27].   That
% is, their assembly must be algorithmic.  Glycans may be considered a
% natural realization of the Wang construct,with monomers acting like
% tiles whose stickiness is encodedby GTase enzymes.  If glycan
% biosynthesis could be madealgorithmic, cells could suppress unwanted
% byproducts andgenerate only desired glycan oligomers with high yield.

\end{document}