### 2020-10-14 papers/sugar-synth/related.tex

% Success of formal methods
Formal methods have successful in applications to a vast range of
problems from the analysis of
systems~\cite{biere1999symbolic2,biere2003bounded,lattice77,ClarkeCEGAR}.
Synthesis of programs that implements a given behavior
has been a focus of research for a while~\cite{PnueliSynthesis}.
In recent years, the synthesis of programs using SAT/SMT solvers
has gained momentum~\cite{SrivastavaSynthesis,Solar-Lezama2005,sygus}.
%
The approach encodes the search of a program that exhibits a certain
behavior into a satisfiability problem.
%
The solvers attempt to find a solution of the satisfiability problem.
%
The solution is the synthesized program.

% Boolean networks

% Program synthesis
Since the formal methods have been exhibiting effectiveness in the vast
range of problems, the
systems biology community has been applying the methods for various
biological problems~\cite{fisher2007executable}.
The key focus has been to model gene regulatory networks (GRNs)~\cite{wang2012BooleanOverview,delayedCTMC}.

There are many biological processes where we can observe the system behavior,
but the exact mechanisms of the processes are not known.
%
Recently, the methods for formal synthesis are finding their applications in biology~\cite{dunn2014defining,xuPluripotency,booleanModelKarp13,paoletti2014analyzing,koksal2013synthesis,fisher2015synthesising}.
%
As far as we know, there has been no similar computation based analysis of glycan production rules.

% In~\cite{},
% GRNs with unknown interactions are modeled using
% a automata-like model with missing components
% and the specification is the variations of the behavior of the system under mutations.
% Their approach uses a counterexample-guided inductive synthesis (CEGIS) based algorithm~\cite{cegis}.
% CEGIS is a framework for synthesis. It first finds a program that may satisfy `samples' from
% the specification. If the synthesized program satisfies the specifications, the CEGIS terminates.
% Otherwise, the method learns a new sample where the program violates the specifications.
% It adds the sample to the set and goes to the next iteration.
% Typically, the constraint solvers find the programs during intermediate steps of CEGIS by
% solving the generated queries.
% %
% Our method follows a similar pattern, where
% a set of sampled observations is a specification, i.e., output molecules,
% and we need to find the governing programs, i.e., production rules.
% %
% In a subsequent work~\cite{}, a Boolean network model is used,
% the functions attached to the nodes are considered unknown, and
% a different kind of data sets provides the behavior specification.
% In this approach, their earlier method is adopted to work in the new situation.
% %

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: