### 2020-10-15 papers/sugar-synth/variations.tex
variations.tex

In the previous section, we have presented a simplified version of
the synthesis problem.
However, the biological problem has more details.
% There are unsettled modeling choices.
We have developed % and implemented
several variations of the
above method.
% A user may choose the variation depending on the choice of model.

\subsection{Compartments}

The production rules may live in different compartments.
The compartments are ordered.
A molecule moves from one compartment to the next.
The rules of the current compartment apply to the molecule.
We have presented the formal description of the compartments in
section~\ref{sec:model}.
To support the compartments, we take an additional integer
input $k$ in $\textsc{SugarSynth}$ to indicate the maximum number
of compartments.
We construct each template rule with a new variable with domain $[1,k]$
indicating the compartment of the rule.
Let $v_r$ be a node of some template rule. We will write $compart(v_r)$
for the compartment variable for the template rule.

We will alter $\textsc{EncodeProduce}$ and its sub-procedures to
ensure that they enforce the compartment order.
Essentially, we need to encode that a rule is applied when all the
pattern nodes were added in the current or earlier compartment.
For the encoding, we add a new map $compart$ that maps
nodes to variables with domain $[1,k]$.
We modify the function $\textsc{MatchTree}$ by replacing
the {\bf if} by the following code.
\begin{algorithmic}[1]
  \vspace{1ex}
  \setcounterref{ALG@line}{line:matchtree-v-absent}
  \If{$isExpand$}
   \State $tCons := ( mark \leq \tau(v)  \land compart(v_r) = compart(v) )$
  \Else
   \State $tCons := ( \tau(v) < mark  \land compart(v_r) \geq compart(v) )$
   \EndIf
   \vspace{1ex}
\end{algorithmic}
In the above code, we have inserted constraints related to compartments
as discussed earlier.
Similarly, we also need to modify the function
$\textsc{EncodeP}$ by inserting the following line after~\ref{line:encodep-ans-match}.
\begin{algorithmic}[1]
  \vspace{1ex}
  \setcounterref{ALG@line}{line:encodep-ans-match}
  \State $c \landplus compart(v_r) \geq compart(v')$
  \vspace{1ex}
\end{algorithmic}
Using the above constraints, we can divide the learned rules
into compartments.

% \subsection{Compartment stay model}

% When molecules pass through compartments, they are expected to stay
% for a while in the compartments.
% %
% The length of the stay determines the number of rule applications
% occurring on the molecule.
% %
% In our method, we assume that the molecules may stay any length of time
% in a compartment, which may result in any number of applications of the rules.
% %
% % This allows the maximum number of molecules are produced for a given set of rules.
% %
% We can assume a whole range of {\em stay models} for the synthesis.
% %
% We have considered one more stay model, where a molecule stays
% in the compartment until no more rule applications are possible.
% %
% We call the stay model as {\em infinite stay}.

% To encode the infinite stay model, we need to add constraints that
% say when a molecule goes to the next compartment, no rule of
% the current compartment is applicable.
% %
% We add the following constraints in \textsc{EncodeProduce}.
% $$
% \Land_{v\in m}\Land_{i \text{ such that } C(v,i) \neq \bot}\Land_{t \in T}
% compart( v ) \leq compart(t) < compart(C(v,i)) 
% \limplies \Land noMatch(C(v,i),t)
% $$
% where constraint $noMatch(v,t)$ is collected in \textsc{EncodeP}
% by inserting the following code after the while loop and an update on $c$ before making the last two updates on $c$.
% $noMatch(v,t)$ is $\ltrue$
% before the call to \textsc{EncodeP}.
% \begin{algorithmic}[1]
%   \vspace{1ex}
%   \setcounterref{ALG@line}{line:encodep-vr-expand}
%   \State $nomatch(v,t) \landplus \lnot c$
%   \vspace{1ex}
% \end{algorithmic}
% The constraints state that for each node $v$ and template rule $t$
% if a rule adds node $v$ before or at the compartment of $t$
% and its $i$th child after the compartment of $t$,
% then $t$ must not be applicable at the $i$th child.
% The above constraints effectively encode that the molecule
% cannot expand at some node until all the relevant rules on
% earlier compartments are disabled.
% Note that if there is no child at the $i$th node, we are adding no constraints of disabling rules of future compartments.
% One may also add those as a requirement, depending on
% the interpretation of the stay model.

\subsection{Fast and slow reactions}
During the stay of a molecule in a compartment, it could be observed that some reactions were dominating than others. This can be explained by characterizing reactions as either slow or fast. A slow reaction can occur only when no other fast reaction is able to extend the molecule in that compartment. We can now define \textsc{Extend} recursively as follows:
\begin{align*}
  \textsc{Extends}(r, m) :=\;&   Apply(m,r) \lor
  \Lor_{\mathclap{i \in [1,w]}} \textsc{Extends}( r, C(m,i ) )\\
    \textsc{Extends}(r, \bot) :=\;&  \nu(v) \neq \bot
  \end{align*}
We now modify $\textsc{EncodeP}$ after~\ref{line:mcut}:
\begin{algorithmic}[1]
  \vspace{1ex}
  \setcounterref{ALG@line}{line:mcut}
  \State $c \landplus \lnot \textsc{Fast}(v_r) \implies 
  \Land_{t \in T} (\textsc{Fast}(t) \implies \lnot\textsc{Extends}(t,v))$
  \vspace{1ex}
\end{algorithmic}
Here, \textsc{Fast} sets the constraint on the rule to be fast.
A negative molecule which is a proper subtree of an input molecule will cease to be negative if there is any fast reaction that is able to extend it as fast reactions happen aggressively and can make partial molecules complete. Therefore, we modify the function $\textsc{SugarSynth}$ after~\ref{line:encode-neg-mol}:
\begin{algorithmic}[1]
  \vspace{1ex}
  \setcounterref{ALG@line}{line:encode-neg-mol}
  \State $nCons \landplus
  \lnot  \Lor_{r \in R} \textsc{Fast}(r) \land \textsc{Extends}(r,m')$
  \vspace{1ex}
\end{algorithmic}
Here, \textsc{Fast} constraints the rule to be fast. The constraints state that none of the rules should be both fast and able to extend the negative molecule received from the model.

\subsection{Unbounded molecules}
We always observe a finite set of molecules. However,
a set of production rules is capable of producing
an unboundedly large size of molecules. In such cases, rules are allowed to form molecules having repeating patterns of a subtree while rest of the tree being exactly same as one of the input molecules. Constraints are added to encode that those rules are all applied in the same compartment and are slow reactions. We put a limit on the depth of such repeating subtrees ($d_0$) and the number of repetitions ($r_0$) . Then, for a particular depth $d$ and repetitions $r$, we recursively define \textsc{Repeat} and \textsc{ExactMatch} as follows:
\begin{align*}
  \textsc{Repeat}(m, m', r, d) :=\;&  \textsc{Repeat}(m, ancestor(m'), r, d-1)\\
    \textsc{Repeat}(m, m', r, 0) :=\;& (compart(m) = compart(m')) \land \textsc{Repeat}(m, C^{2d}(m'), r-1, d) \\
    \textsc{Repeat}(m, m', 0, d) :=\;&  \textsc{ExactMatch}(m,C^{2d}(m')) \\
    \textsc{Repeat}(\bot, \bot, r, d) :=\;&  \top \\
    \textsc{Repeat}(m, \bot, r, d) :=\;&  \bot \\
    \textsc{Repeat}(\bot, m', r, d) :=\;&  \bot \\
    \textsc{ExactMatch}(m,m') :=\;& (M(m) = M(m')) \land \Land_{i \in [1,w]} \textsc{ExactMatch}(C(m,i),C(m',i)) \\
    \textsc{ExactMatch}(\bot,m') :=\;& \bot \\
    \textsc{ExactMatch}(m,\bot) :=\;& \bot 
  \end{align*}
  where $ancestor(m')$ is the parent node of $m'$ and $C^{2*d}$ is the application of $C(m',i)$, $2d$ times and $i$ comes from the reverse of the path traversed by m' to reach it's $d^{th}$ parent.
The constraints for allowing runaway reactions are added after~\ref{line:consNewR} to modify \textsc{SugarSynth}.
\begin{algorithmic}[1]
  \vspace{1ex}
  \setcounterref{ALG@line}{line:consNewR}
  \State $mCons \landplus \Lor_{1 \leq r \leq r_0} \Lor_{1 \leq d \leq d_0} \Lor_{m \in \mu} (M(m) \neq M(\hat{m})) \implies  \textsc{Repeat}(m, \hat{m}, r, d) $
  \vspace{1ex}
\end{algorithmic}
where \textsc{Repeat} adds the constraints described above due to which runaway reactions are allowed.

\subsection{Non-monotonic rules}
Some molecules produced dictate that after a certain rule is applied to them in a compartment, it restricts the application of any other rule from that point of time. Such a rule is said to have \textsc{HardEnds}. This feature of rules help in producing an exact set of desired molecules. We modify the function $\textsc{MatchTree}$ by inserting the following constraints after~\ref{line:matchtree-cons}:
\begin{algorithmic}[1]
  \vspace{1ex}
  \setcounterref{ALG@line}{line:matchtree-cons}
  \State $c \landplus \textsc{HardEnds}(v_r) \implies (mark \leq \tau(v))$
  \vspace{1ex}
\end{algorithmic}
In our method, \textsc{HardEnds} sets a constraint on the given rule to have hard ends. The constraints state that if the applicable rule has \textsc{HardEnds}, then it has to be added at a time earlier than the current time of the molecule, effectively restricting the addition of further rules.

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: