### 2021-07-04 papers/sugar-synth/main.tex

\documentclass{llncs}
%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
% \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

% \documentclass[conference,10pt]{IEEEtran}

% \bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
% \citestyle{acmauthoryear}   %% For author/year citations

\input{macro.tex}
\setlength{\textfloatsep}{0pt}%

\title{Automated inference of production rules for glycans}
\author{Ansuman Biswas$^2$, Ashutosh Gupta$^1$, Meghana Missula$^1$, and Mukund Thattai$^2$}
\institute{$^1$IITB and $^2$NCBS}

\date{May 2019}

\begin{document}

\maketitle

\begin{abstract}
  Glycans are tree-like polymers made up of sugar monomer building blocks. They are found on the surface of all living cells, and distinct glycan trees act as identity markers for distinct cell types. Proteins called GTase enzymes assemble glycans via the successive addition of monomer building blocks. The rules by which the enzymes operate are not fully understood. In this paper, we present the first SMT-solver-based iterative method that infers the assembly process of the glycans by analyzing the set of glycans from a cell. We have built a tool based on the method and applied it to infer rules based on published glycan data.
\end{abstract}

\section{Introduction}
\label{sec:intro}
\input{intro}

\section{Production of Glycans}
\label{sec:bio}
\input{bio.tex}

% \clearpage

\section{Motivating example}
\label{sec:motivation}
\input{motivation.tex}

\section{Modelling of the synthesis problem}
\label{sec:model}
\input{model.tex}

% \clearpage

\section{Method for the synthesis problem}
\label{sec:algo}
\input{algo.tex}

\section{Experiments}
\label{sec:experiments}
\input{experiments.tex}

% \section{Related work}
% \label{sec:related}
% \input{related.tex}

\section{Conclusion and future work}
% ##limitations in the model because of limited information,
% a)the boundary between two compartments is fuzzy, 
% b) the growths of two branches are correlated/ not correlated..and some more points like these, for which wealth of data is not available, hence the model was unable to incorporate the specifics
\label{sec:conclusion}
\input{conclusion.tex}
\nocite{Charles2013}
\bibliographystyle{unsrt}
\bibliography{biblio}

\appendix

% 
% To resolve this paradox we borrow from the field of
% al-gorithmic self-assembly, which explores how small buildingblocks
% with stochastic local interactions assemble into various global
% configurations [23]–[25].  A central inverse problemin self-assembly
% is to design building blocks that assembleinto a target shape.  This
% process has been studied using thetheoretical framework of Wang tiles:
% square building blockswith colored sides that stick to one another
% along sides withmatching colors [26].  For efficient assembly of Wang
% tiles,their coloring must be cleverly chosen so the target shape isthe
% unique terminal output of the stochastic assembly process[27].   That
% is, their assembly must be algorithmic.  Glycans may be considered a
% natural realization of the Wang construct,with monomers acting like
% tiles whose stickiness is encodedby GTase enzymes.  If glycan
% biosynthesis could be madealgorithmic, cells could suppress unwanted
% byproducts andgenerate only desired glycan oligomers with high yield.

\end{document}

%--------------------- DO NOT ERASE BELOW THIS LINE --------------------------

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: